---
title: "DayMet_Extraction"
format: html
editor: source
---

#loading libraries
```{r}
#remotes::install_github("ropensci/USAboundaries")
#remotes::install_github("ropensci/USAboundariesData")

library(tidyverse)
library(USAboundaries) # for US state boundaries
library(sf) # for US map
library(daymetr)
library(purrr)
```

#MAKE MERGED TRAINING FILE 
```{r}
training_metadata1 <- read_csv("../data/training_metadata1.csv")

training_metadata1
```

##Downloading weather data
```{r}
training_metadata1
daymet_one <- download_daymet(site = training_metadata1$site[1],
                              lat = training_metadata1$latitude[1],
                              lon = training_metadata1$longitude[1],
                              start = training_metadata1$year[1],
                              end = training_metadata1$year[1],
                              simplify = T,
                              silent = T
                              )
daymet_one
```

```{r}
daymet_all <- training_metadata1 %>%
  mutate(weather = pmap(list(.y = year,
                             .site = site,
                             .lon = longitude,
                             .lat = latitude),
                        function(.y, .site, .lat, .lon)
                          download_daymet(
                                          site = .site,
                                          lat = .lat,
                                          lon = .lon,
                                          start = .y,
                                          end = .y,
                                          simplify = T,
                                          silent = T) %>%
                          rename(.year = year,
                                 .site = site)
                          
                        ))
daymet_all
```


##Inspecting weather data for the first site-year. 
```{r}
daymet_all$weather[1]

```


##Unnesting the weather column and making Growing Degree Day column 
```{r}
daymet_all_unnest1 <- daymet_all %>%
  mutate(weather = map(weather, ~ select(.x, -latitude, -longitude))) %>%
  unnest(weather) %>%
  pivot_wider(
    names_from = measurement,
    values_from = value
  ) %>%
  janitor::clean_names() %>%
  mutate(tmax30_deg_c = pmin(tmax_deg_c, 30)) %>% #above 30, temp does not increase GDD
  mutate(tmin10_deg_c = pmax(tmin_deg_c, 10)) %>%
  mutate(gdd = round((tmax30_deg_c + tmin10_deg_c)/2) - 10) %>% #round any value above 30 to 30 and calculate growing degree days for corn and round decimals to only 2 places
  mutate(gdd_rounded = pmax(gdd, 0)) %>%
  select(-site_2, -year_2)

daymet_all_unnest1
```


##Exporting file for future use.  
```{r}
write_csv(daymet_all_unnest1,
          "../data/fieldweatherdata.csv")
```

#MAKE MERGED SUBMISSION FILE 
```{r}
testing_meta <- read_csv("../data/testing/testing_meta.csv")

testing_meta
```

##Downloading weather data
```{r}
testing_meta
daymet_one_testing <- download_daymet(site = testing_meta$site[1],
                              lat = testing_meta$latitude[1],
                              lon = testing_meta$longitude[1],
                              start = testing_meta$year[1],
                              end = testing_meta$year[1],
                              simplify = T,
                              silent = T
                              )
daymet_one_testing
```

```{r}
daymet_all_testing <- testing_meta %>%
  mutate(weather = pmap(list(.y = year,
                             .site = site,
                             .lon = longitude,
                             .lat = latitude),
                        function(.y, .site, .lat, .lon)
                          download_daymet(
                                          site = .site,
                                          lat = .lat,
                                          lon = .lon,
                                          start = .y,
                                          end = .y,
                                          simplify = T,
                                          silent = T) %>%
                          rename(.year = year,
                                 .site = site)
                          
                        ))
daymet_all_testing
```


Inspecting weather data for the first site-year. 
```{r}
daymet_all_testing$weather[1]

```


#Unnesting the weather column and making Growing Degree Day column 
```{r}
daymet_all_unnested_testing <- daymet_all_testing %>%
  mutate(weather = map(weather, ~ select(.x, -latitude, -longitude))) %>%
  unnest(weather) %>%
  pivot_wider(
    names_from = measurement,
    values_from = value
  ) %>%
  janitor::clean_names() %>%
  mutate(tmax30_deg_c = pmin(tmax_deg_c, 30)) %>% #above 30, temp does not increase GDD
  mutate(tmin10_deg_c = pmax(tmin_deg_c, 10)) %>%
  mutate(gdd = round((tmax30_deg_c + tmin10_deg_c)/2) - 10) %>% #round any value above 30 to 30 and calculate growing degree days for corn and round decimals to only 2 places
  mutate(gdd_rounded = pmax(gdd, 0)) %>%
  select(-site_2, -year_2)

daymet_all_unnested_testing 
```


#Exporting file for future use.  
```{r}
write_csv(daymet_all_unnested_testing,
          "../data/fieldweatherdata_testing.csv")
```