---
title: "DayMet_Extraction"
format: html
editor: source
---

#loading libraries
```{r}
#remotes::install_github("ropensci/USAboundaries")
#remotes::install_github("ropensci/USAboundariesData")

library(tidyverse)
library(USAboundaries) # for US state boundaries
library(sf) # for US map
library(daymetr)
library(purrr)
```

<<<<<<< HEAD
```{r}
training_metadata1 <- read_csv("../data/training_metadata1.csv")

training_metadata1

=======

#Downloading weather data
```{r}
training_metadata1
>>>>>>> e29ffd977e1c870a98f3df0565f91f82793a362a
daymet_one <- download_daymet(site = training_metadata1$site[1],
                              lat = training_metadata1$latitude[1],
                              lon = training_metadata1$longitude[1],
                              start = training_metadata1$year[1],
                              end = training_metadata1$year[1],
<<<<<<< HEAD
                              simplify = T
=======
                              simplify = T,
                              silent = T
>>>>>>> e29ffd977e1c870a98f3df0565f91f82793a362a
                              )
daymet_one
```

<<<<<<< HEAD
=======

#Downloading weather data
>>>>>>> e29ffd977e1c870a98f3df0565f91f82793a362a
```{r}
daymet_all <- training_metadata1 %>%
  mutate(weather = pmap(list(.y = year,
                             .site = site,
                             .lon = longitude,
                             .lat = latitude),
                        function(.y, .site, .lat, .lon)
                          download_daymet(
                                          site = .site,
                                          lat = .lat,
                                          lon = .lon,
                                          start = .y,
                                          end = .y,
                                          simplify = T,
                                          silent = T) %>%
                          rename(.year = year,
                                 .site = site)
                          
                        ))
daymet_all
```


Inspecting weather data for the first site-year. 
```{r}
daymet_all$weather[1]

```


Unnesting the weather column.  
```{r}
daymet_all_unnest <- daymet_all %>% unnest(weather, names_repair = "unique") %>% #prevents duplication of column names
  pivot_wider(names_from = measurement,
              values_from = value) %>%
  janitor::clean_names()


daymet_all_unnest
```

Alternatively, we can drop longitude and latitude instead of using unique
```{r}
daymet_all_unnest1 <- daymet_all %>%
  mutate(weather = map(weather, ~ select(.x, -latitude, -longitude))) %>%
  unnest(weather) %>%
  pivot_wider(
    names_from = measurement,
    values_from = value
  ) %>%
  janitor::clean_names() %>%
  mutate(tmax_deg_c_rounded_down = pmin(tmax_deg_c, 30)) %>% #round any value above 30 to 30
  mutate(gdd = ((tmax_deg_c_rounded_down - tmin_deg_c)/2) - 10) %>% #calculate growing degree days for corn
  round(gdd, 2) #round decimals to only 2 places
  


daymet_all_unnest1
```
###Notes:
Must round any Tmax value down to 30C if it exceeds 30
Must reduce GDD to just 2 decimal points

Exporting file for future use.  
```{r}
write_csv(daymet_all_unnest1,
          "../data/fieldweatherdata.csv")
```


```{r}

daymet_all$weather[1]  #inspect weather data for the first site year

daymet_all_unnest <- daymet_all %>%  #unnest weather data to get the Tmin and Tmax
  unnest(weather) %>%
  pivot_wider(names_from = measurement,
              values_from = value) %>%
  janitor::clean_names() %>%
  


daymet_all_unnest
```
