---
title: "Data Science Applied to Ag - Final Project - ML"
format:
  html:
    embed-resources: true
    toc: true
    theme: cerulean
author: Group 1 (Mari, Oluwatomide)
---
#1. Loading Library
```{r, warning = False, message = False}
library(tidyverse)
library(randomForest)
library(xgboost)
library(caret)
library(lubridate)
library(readr)
library(dplyr)
library(tidymodels)
library(vip)
library(ranger)
library(finetune)
library(ggridges)
library(doParallel)
```



#2. Viewing Files
```{r, warning = False, message = False}
training_metadata <- read_csv("../data/training/training_meta.csv")

training_soil <- read_csv("../data/training/training_soil.csv")

training_trait <- read_csv("../data/training/training_trait.csv")

testing_submission <- read_csv("../data/testing/testing_submission.csv")

#view(training_metadata)
#view(training_soil)
#view(training_trait)
view(testing_submission)

```


#3. Removing Duplicates and determining day of the year
```{r, warning = False, message = False}

training_soil1 <- training_soil %>%
  select(!(year)) %>%
  separate(site,
                   into = c("site", "year")) %>%
  mutate(year = as.numeric(year)) %>%
  filter(nchar(as.character(year)) == 4) 

training_metadata1<- training_metadata %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>% #remove x column
  unique() %>%
  filter(
    latitude >= 24.5 & latitude <= 49.5,
    longitude >= -125 & longitude <= -66.5
  ) ##Removing sites outside US boundaries

write_csv(training_metadata1, "../data/training_metadata1.csv")

training_trait1 <- training_trait %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>% #remove x column with the -Dry, -Early, -Late
  mutate(
    across(c(date_planted, date_harvested),
           ~as.Date(.x, format = "%m/%d/%y")),
    
    day_of_year_planted = yday(date_planted),
    day_of_year_harvested = yday(date_harvested), .before = date_harvested
  )

view(training_soil1)
view(training_metadata1)
view(training_trait1)
```

#4. Creating Merged Dataset
```{r, warning = False, message = False}
training_merged <- training_trait1 %>%
  left_join(training_soil1, by = c("site", "year")) %>%
  left_join(training_metadata1, by = c("site", "year"))

view(training_merged)
```

#5. Mapping Sites
```{r, warning = False, message = False}
library(USAboundaries)
library(ggplot2)
library(dplyr)
library(sf)
site_locations <- us_states() %>%
filter(!(state_abbr %in% c("PR", "AK", "HI")))
ggplot() +
  geom_sf(data = site_locations) +
  geom_point(data = training_merged, 
             aes(x = longitude,
                 y = latitude,
                color = yield_mg_ha)
  )

#Making a map to identify the site distributions. PR, AK, HI were removed since they did not have sites. Points exist in the Mediterranean between Corsica and Italy, i.e. Site NEH2 (which exists in the final testing_submission file).
```


#6. Matching Sites Across Files
```{r, warning = False, message = False}

#Making Sites Only Identical Across Training and Testing

setdiff(testing_submission$site, training_merged$site)  # sites in df1 but not df2
setdiff(training_merged$site, testing_submission$site)  # sites in df2 but not df1

#Drop sites from training_merged that to not need to be predicted to complete testing_submission. ONH1 and 2 are in proximity to one another, however to not appear in the testing_submission dataframe. ONH3 (a new site) appears in testing_submission. Potential to predict ONH3 based on ONH1 and ONH2, so will not remove them.

training_merged1 <- training_merged %>%
  filter(!(site %in% c("IAH1a" ,"IAH1b" ,"IAH1c", "IAH3" , "MOH2" , "NYH1" ,"TXH2" , "IAH1" , "KSH1" , "NEH4" , "SDH1" , "ARH1" , "ARH2",  "COH1" ,"GEH1",  "TXH3" , "TXH4" , "NYS1" ))) %>%
  mutate(id = row_number(), .before = "year") #adding an ID column at the front of the table to identify rows

write_csv(training_merged1,"../data/training_merged1.csv")

#confirm: 
setdiff(testing_submission$site, training_merged1$site)  # sites in df1 but not df2
setdiff(training_merged1$site, testing_submission$site)  # sites in df2 but not df1

view(training_merged1)
```

```{r, warning = False, message = False}
fieldweather <- read_csv("../data/fieldweatherdata.csv")

fieldweatherdata <- fieldweather %>%
  select(-gdd)

summary(fieldweatherdata)

```


#7. Density plots  
```{r}
#To explore the weather data distributions

fieldweatherdata %>% 
  pivot_longer(cols = dayl_s:gdd_rounded) %>%
  ggplot(aes(x = value)) +
  geom_density() + 
  facet_wrap(~name, scales = "free")

```

```{r}
fieldweather1 <- fieldweatherdata %>%
  # Selecting needed variables
  dplyr::select(-tile, -altitude) %>%
# Creating a date class variable  
  mutate(date_chr = paste0(year, "/", yday)) %>% #paste0 helps to merge columns together
  mutate(date = as.Date(date_chr, "%Y/%j"))

```

Merging file with weather variables by growing window (date planted:date harvested)
```{r}
weather_merged <- training_merged1 %>%
  rowwise() %>%
  mutate(
    weather_summary = list(
      fieldweather1 %>%
        filter(date >= date_planted,
               date <= date_harvested) %>%
        summarise(across(
            .cols = c(dayl_s, srad_w_m_2, tmax_deg_c, tmin_deg_c, vp_pa),
            .fns = mean,
            .names = "mean_{.col}"),
          across(
            .cols = c(gdd_rounded, prcp_mm_day),
            .fns = sum,
            .names = "sum_{.col}"
          ))))%>%
  unnest(weather_summary) %>%
  ungroup %>% 
  select(-id, -replicate, -block)

head(weather_merged)

write_csv(weather_merged, "../output/weather_merged.csv")

#######SUM GDD ROUNDED IS 10 ORDERS TOO LARGE#######
```


# 8. Data Visualization 
```{r}
#Plotting all variables by site

allplots <- weather_merged %>%
  pivot_longer(mean_dayl_s:sum_prcp_mm_day) %>%
  group_by(name) %>%
  nest() %>%
  mutate(plot = map2(data, name,
                     ~ggplot(data = .x,
       aes(x = value,
           y = site,
           fill = stat(x))
       ) +
  geom_density_ridges_gradient(scale = 3,
                               rel_min_height = 0.01) +
  scale_fill_viridis_c(option = "C", alpha = .8) +
  theme(legend.position = "none") +
  labs(x= .y,
       y = "site")))
  
allplots  
```

```{r}
allplots$plot
```

#9. Pre-Processing Before MLA
```{r}
set.seed(76332)

# Splitting the Data for Training
merged_split <- initial_split(weather_merged, prop = 0.7, strata = yield_mg_ha)
merged_train <- training(merged_split) %>%
  select(-c(date_planted, date_harvested))
merged_test <- testing(merged_split)

```

10. Distribution of Yield
```{r}
ggplot() +
  geom_density(data = merged_train, 
               aes(x = yield_mg_ha),
               color = "red") +
  geom_density(data = merged_test, 
               aes(x = yield_mg_ha),
               color = "blue") 
```

11. Recipe and Prep for Training Model
```{r}
#RECIPE
merged_recipe <- recipe(yield_mg_ha ~ ., data = merged_train) %>%
  step_novel(all_nominal_predictors()) %>% #renames "new" hybrids as "new" category so that the model does not return an error
  step_unknown(all_nominal_predictors()) %>% #handling NAs in categorical columns
  step_dummy(all_nominal_predictors()) %>%  #make dummy variables on categorical variables like previous crop
  step_impute_median(all_numeric_predictors())#imputing NA values with median

#DATA PREP
merged_prep <- merged_recipe %>%
  prep()

merged_prep
```

12. Random Forest
```{r}
library(tidymodels)
library(tidyverse)
library(glmnet)
library(vip)
library(ranger)
library(finetune)
```

Specifying rf as our model type, asking to tune the hyperparameters
```{r}
rf_spec <- rand_forest(trees = tune(),
                       mtry = tune()) %>%
  set_engine("ranger") %>%
  set_mode("regression")
rf_spec
```


Defining our resampling strategy, using a 20-fold cross validation approach:  
```{r resampling_foldcv}
set.seed(134549)
resampling_foldcv <- vfold_cv(merged_train, 
                              v = 20)

resampling_foldcv
resampling_foldcv$splits[[1]]
```


Restricting iterations to 100 for our algorithm
```{r rf_grid_result}
set.seed(1276544)
rf_grid_result <- tune_sim_anneal(object = rf_spec,
                     preprocessor = merged_recipe,
                     resamples = resampling_foldcv,
                    iter = 100
                     )

beepr::beep()

rf_grid_result
rf_grid_result$.metrics[[2]]
```


Calculating RMSE
```{r RMSE}
rf_grid_result %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  ggplot(aes(x = mtry, 
             y = trees 
             )) +
  geom_path(group = 1) +
  geom_point(aes(color = mean),
             size = 3) + 
  scale_color_viridis_b() +
  geom_text(aes(label = .iter), nudge_x = .0005) +
  labs(title = "RMSE")
```

Calculating R2
```{r R2}
rf_grid_result %>%
  collect_metrics() %>%
  filter(.metric == "rsq") %>%
  ggplot(aes(x = mtry, 
             y = trees 
             )) +
  geom_path(group = 1) +
  geom_point(aes(color = mean),
             size = 3) + 
  scale_color_viridis_b() +
  geom_text(aes(label = .iter), nudge_x = .0005) +
  labs(title = "R2")

```


Selecting the best model based on RMSE (lowest) 
```{r}
best_rf_rmse <- rf_grid_result %>%
  select_by_pct_loss("trees", 
                     metric = "rmse",
                     limit = 2)

best_rf_rmse

```

Selecting the best model based on R2 (highest) 
```{r}
best_rf_r2 <- rf_grid_result %>%
  select_by_pct_loss("trees",
                     metric = "rsq",
                     limit = 2)
best_rf_r2

```

Let's use the hyperparameter values that optimized R2 to fit our final model.

```{r final_spec}
final_rf_spec <- rand_forest(trees = best_r2$trees,
                          mtry = best_r2$mtry) %>%
  set_engine("ranger",
             importance = "permutation") %>%
  set_mode("regression")
  

final_rf_spec
```


Validation
```{r}
final_rf_fit <- last_fit(final_rf_spec,
                merged_recipe,
                split = merged_split)

final_rf_fit %>%
  collect_predictions()
```

Metrics on the **test set**:
```{r}
final_rf_fit %>%
  collect_metrics()
```

Metrics on **train set** 
```{r}
# RMSE
final_rf_spec %>%
  fit(yield_mg_ha ~ .,
      data = bake(merged_prep, 
                  merged_train)) %>%
  augment(new_data = bake(merged_prep, 
                          merged_train)) %>% 
  rmse(yield_mg_ha, .pred) %>%
  bind_rows(
    
    
    # R2
    final_rf_spec %>%
      fit(yield_mg_ha ~ .,
          data = bake(merged_prep, 
                      merged_train)) %>%
      augment(new_data = bake(merged_prep, 
                              merged_train)) %>% 
      rsq(yield_mg_ha, .pred)
    
  )

```

Predicted vs. observed plot:  
```{r}
final_rf_fit %>%
  collect_predictions() %>%
  ggplot(aes(x = yield_mg_ha,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

Variable of importance
```{r}
final_rf_spec %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, weather_merged)) %>%
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
    
```


13. Elastic Net
```{r}
enet_spec <-
  # Specifying linear regression as our model type, asking to tune the hyperparameters
linear_reg(penalty = tune(),
           mixture = tune()
           ) %>%
  # Specify the engine
set_engine("glmnet")

enet_spec
```

Hyper-parameter tuning  
Creating a **grid** to perform our hyperparameter tuninig search with multiple combinations of values for penalty and mixture:

```{r enet_grid}
enet_grid <- crossing(penalty = seq(0, 
                                    10, 
                                    by = 1),
                      mixture = seq(0, 
                                    1, 
                                    by = 0.1))

enet_grid

ggplot(data = enet_grid,
       aes(x = mixture,
           y = penalty)) +
  geom_point()
```

Grid search 
```{r enet_grid_result}
enet_grid_result <- tune_grid(enet_spec,
                              preprocessor = merged_recipe,
                              grid = enet_grid,
                              resamples = resampling_foldcv)

enet_grid_result
enet_grid_result$.metrics[[1]]
```

RMSE for Enet
```{r RMSE}
enet_grid_result %>%
  collect_metrics() %>% 
  filter(.metric == "rmse") %>%
  ggplot(aes(x = penalty, 
             y = mean, 
             color = factor(mixture), 
             group = factor(mixture))) +
  geom_line() +
  geom_point() + 
  labs(y = "RMSE")
```

R2 for Enet
```{r R2}
enet_grid_result %>%
  collect_metrics() %>%
  filter(.metric == "rsq") %>%
  ggplot(aes(x = penalty, 
             y = mean, 
             color = factor(mixture), 
             group = factor(mixture))) +
  geom_line() +
  geom_point() + 
  labs(y = "R2")
```

Extracting hyperparameters from the best model based on lowest RMSE
```{r}
best_enet_rmse <- enet_grid_result %>% select_best(metric = "rmse")

best_enet_rmse
```

Extracting hyperparameters from the best model based on highest R2
```{r}
best_enet_r2 <- enet_grid_result %>%
  select_best(metric = "rsq")

best_enet_r2
```

Let's use the hyperparameter values that optimized R2 to fit our final model.
```{r final_spec}
final_enet_spec <- linear_reg(penalty = 0.7,
                         mixture = 0) %>%
  set_engine("glmnet")

final_enet_spec
```


```{r final_fit}
final_enet_fit <- last_fit(final_enet_spec,
                      merged_recipe,
                      split = merged_split)

final_enet_fit %>%
  collect_predictions()
```

Metrics on the **test set**:
```{r}
final_enet_fit %>% collect_metrics()
```

Metrics on **train set** 
```{r}
# RMSE
final_enet_spec %>%
  fit(yield_mg_ha ~ .,
      data = bake(merged_prep, 
                  merged_train)) %>%
  augment(new_data = bake(merged_prep, 
                          merged_train)) %>% 
  rmse(yield_mg_ha, .pred) %>%
  bind_rows(
    
    
    # R2
    final_enet_spec %>%
      fit(yield_mg_ha ~ .,
          data = bake(merged_prep, 
                      merged_train)) %>%
      augment(new_data = bake(merged_prep, 
                              merged_train)) %>% 
      rsq(yield_mg_ha, .pred)
    
  )

```

Predicted vs. observed plot:  
```{r}
final_enet_fit %>%
  collect_predictions() %>%
  ggplot(aes(x = yield_mg_ha,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

Coefficients:  
```{r}
final_enet_spec %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, weather_merged)) %>%
  tidy() %>%
  arrange(desc(estimate))

```

Variable of importance
```{r}
final_enet_spec %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, weather_merged)) %>%
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
    
```



14. XGBoost
```{r}

```




16. Comparison of RMSE and R2 Values
```{r}
#Make a table of the RMSE and R2 Values to compare directly
```

17. Final Conclusion
```{r}
#Which model we chose
```
