---
title: "Data Science Applied to Ag - Final Project - ML"
format:
  html:
    embed-resources: true
    toc: true
    theme: cerulean
author: Group 1 (Mari, Oluwatomide)
---
#1. Loading Library
```{r, warning = False, message = False}
#install.packages("lightgbm")
library(tidyverse)
library(lightgbm)
library(xgboost)
library(caret)
library(lubridate)
library(readr)
library(dplyr)
library(tidymodels)
library(vip)
library(ranger)
library(finetune)
library(ggridges)
library(doParallel)
library(glmnet)
library(finetune)
library(janitor)
```



#2. Viewing Files
```{r, warning = False, message = False}
training_metadata <- read_csv("../data/training/training_meta.csv")

training_soil <- read_csv("../data/training/training_soil.csv")

training_trait <- read_csv("../data/training/training_trait.csv")

testing_submission <- read_csv("../data/testing/testing_submission.csv")


```


#3. Removing Duplicates and determining day of the year
```{r, warning = False, message = False}

training_soil1 <- training_soil %>%
  select(!(year)) %>%
  separate(site,
                   into = c("site", "year")) %>%
  mutate(year = as.numeric(year)) %>%
  filter(nchar(as.character(year)) == 4) 

training_metadata1<- training_metadata %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>% #remove x column
  unique() %>%
  filter(
    latitude >= 24.5 & latitude <= 49.5,
    longitude >= -125 & longitude <= -66.5
  ) ##Removing sites outside US boundaries

write_csv(training_metadata1, "../data/training_metadata1.csv")

training_trait1 <- training_trait %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>% #remove x column with the -Dry, -Early, -Late
  mutate(
    across(c(date_planted, date_harvested),
           ~as.Date(.x, format = "%m/%d/%y")),
    
    day_of_year_planted = yday(date_planted),
    day_of_year_harvested = yday(date_harvested), .before = date_harvested
  )

view(training_soil1)
view(training_metadata1)
view(training_trait1)
```

#4. Creating Merged Dataset
```{r, warning = False, message = False}
training_merged <- training_trait1 %>%
  left_join(training_soil1, by = c("site", "year")) %>%
  left_join(training_metadata1, by = c("site", "year"))

view(training_merged)
```

#5. Mapping Sites
```{r, warning = False, message = False}
library(USAboundaries)
library(ggplot2)
library(dplyr)
library(sf)
site_locations <- us_states() %>%
filter(!(state_abbr %in% c("PR", "AK", "HI")))
ggplot() +
  geom_sf(data = site_locations) +
  geom_point(data = training_merged, 
             aes(x = longitude,
                 y = latitude,
                color = yield_mg_ha)
  )

#Making a map to identify the site distributions. PR, AK, HI were removed since they did not have sites. Points exist in the Mediterranean between Corsica and Italy, i.e. Site NEH2 (which exists in the final testing_submission file).
```


#6. Matching Sites Across Files
```{r, warning = False, message = False}

#Making Sites Only Identical Across Training and Testing

setdiff(testing_submission$site, training_merged$site)  # sites in df1 but not df2
setdiff(training_merged$site, testing_submission$site)  # sites in df2 but not df1

#Drop sites from training_merged that to not need to be predicted to complete testing_submission. ONH1 and 2 are in proximity to one another, however to not appear in the testing_submission dataframe. ONH3 (a new site) appears in testing_submission. Potential to predict ONH3 based on ONH1 and ONH2, so will not remove them.

training_merged1 <- training_merged %>%
  filter(!(site %in% c("IAH1a" ,"IAH1b" ,"IAH1c", "IAH3" , "MOH2" , "NYH1" ,"TXH2" , "IAH1" , "KSH1" , "NEH4" , "SDH1" , "ARH1" , "ARH2",  "COH1" ,"GEH1",  "TXH3" , "TXH4" , "NYS1" ))) %>%
  mutate(id = row_number(), .before = "year") #adding an ID column at the front of the table to identify rows

write_csv(training_merged1,"../data/training_merged1.csv")

#confirm: 
setdiff(testing_submission$site, training_merged1$site)  # sites in df1 but not df2
setdiff(training_merged1$site, testing_submission$site)  # sites in df2 but not df1

```

```{r, warning = False, message = False}
fieldweather <- read_csv("../data/fieldweatherdata.csv")

fieldweatherdata <- fieldweather %>%
  select(-gdd)

summary(fieldweatherdata)

```


#7. Density plots  
```{r}
#To explore the weather data distributions

fieldweatherdata %>% 
  pivot_longer(cols = dayl_s:gdd_rounded) %>%
  ggplot(aes(x = value)) +
  geom_density() + 
  facet_wrap(~name, scales = "free")

```

```{r}
fieldweather1 <- fieldweatherdata %>%
  # Selecting needed variables
  dplyr::select(-tile, -altitude) %>%
# Creating a date class variable  
  mutate(date_chr = paste0(year, "/", yday)) %>% #paste0 helps to merge columns together
  mutate(date = as.Date(date_chr, "%Y/%j"))

```


```{r}
weather_merged_new <- training_merged1 %>%
  rowwise() %>%
  mutate(
    weather_summary = list(
      fieldweather1 %>%
        filter(
          site == cur_data()$site,          # match site
          year == cur_data()$year,          #use your year column
          date >= date_planted,
          date <= date_harvested
        ) %>%
        group_by(site, year) %>%
        summarise(
          across(
            .cols = c(dayl_s, srad_w_m_2, tmax_deg_c, tmin_deg_c, vp_pa),
            .fns = mean,
            .names = "mean_{.col}"
          ),
          across(
            .cols = c(gdd_rounded, prcp_mm_day),
            .fns = sum,
            .names = "sum_{.col}"
          ),
          .groups = "drop"
        )
    )
  ) %>%
  unnest(weather_summary) %>%
  ungroup() %>% 
  select(-id, -replicate, -block)

head(weather_merged_new)

write_csv(weather_merged_new, "../output/weather_merged_new.csv")

```



Merging file with weather variables by growing window (date planted:date harvested)
```{r}
weather_merged <- training_merged1 %>%
  rowwise() %>%
  mutate(
    weather_summary = list(
      fieldweather1 %>%
        filter(date >= date_planted,
               date <= date_harvested,
               year(date) == year(date_planted)) %>%#filtering the data based on growing season
        group_by(site, year)%>%
        summarise(across(
            .cols = c(dayl_s, srad_w_m_2, tmax_deg_c, tmin_deg_c, vp_pa),
            .fns = mean,
            .names = "mean_{.col}"),
          across(
            .cols = c(gdd_rounded, prcp_mm_day),
            .fns = sum,
            .names = "sum_{.col}"
          ))))%>%
  unnest(weather_summary) %>%
  ungroup %>% 
  select(-id, -replicate, -block)

head(weather_merged)

write_csv(weather_merged, "../output/weather_merged.csv")

#######SUM GDD ROUNDED IS 10 ORDERS TOO LARGE#######
#######Making loop with growing degree days that iterates by year. #######Currently it is adding from the first year's planting date to the last year's harvest date
```


```{r}
setDT(training_merged1)
setDT(fieldweather1)

training_merged1 <- training_merged1[year != 2014]

#create end of year
training_merged1[, end_of_year := as.Date(paste0(year(date_planted), "-12-31"))]
training_merged1[, harvest_cutoff := pmin(date_harvested, end_of_year)]


weather_merged <- training_merged1 %>%
  filter(year != 2014) %>%
  mutate(
    end_of_year = as.Date(paste0(year(date_planted), "-12-31")),
    harvest_cutoff = pmin(date_harvested, end_of_year)
  ) %>%
  inner_join(fieldweather1, by = "site") %>%
  filter(
    date >= date_planted,
    date <= harvest_cutoff
  ) %>%
  group_by(site, year = year(date_planted), month = month(date)) %>%
  summarise(
    across(
      c(dayl_s, srad_w_m_2, tmax_deg_c, tmin_deg_c, vp_pa),
      ~ mean(.x, na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    across(
      c(gdd_rounded, prcp_mm_day),
      ~ sum(.x, na.rm = TRUE),
      .names = "sum_{.col}"
    ),
    .groups = "drop"
  )

```




#Unique Planting Day - Harvest Day Pairs
```{r}
unique_day_pairs <- weather_merged %>%
  distinct(date_planted, date_harvested)
```



# 8. Data Visualization 
```{r}
#Plotting all variables by site

allplots <- weather_merged %>%
  pivot_longer(mean_dayl_s:sum_prcp_mm_day) %>%
  group_by(name) %>%
  nest() %>%
  mutate(plot = map2(data, name,
                     ~ggplot(data = .x,
       aes(x = value,
           y = site,
           fill = stat(x))
       ) +
  geom_density_ridges_gradient(scale = 3,
                               rel_min_height = 0.01) +
  scale_fill_viridis_c(option = "C", alpha = .8) +
  theme(legend.position = "none") +
  labs(x= .y,
       y = "site")))
  
allplots  
```

```{r}
allplots$plot
```

#9. Pre-Processing Before MLA
```{r}
set.seed(76332)

# Splitting the Data for Training
merged_split <- initial_split(weather_merged, prop = 0.7, strata = yield_mg_ha)

merged_train <- training(merged_split) %>%
  select(-c(date_planted, date_harvested))

merged_test <- testing(merged_split)

```

#10. Distribution of Yield
```{r}
ggplot() +
  geom_density(data = merged_train, 
               aes(x = yield_mg_ha),
               color = "red") +
  geom_density(data = merged_test, 
               aes(x = yield_mg_ha),
               color = "blue") 
```

#11. Recipe and Prep for Training Model
```{r}
#RECIPE
merged_recipe <- recipe(yield_mg_ha ~ ., data = merged_train) %>%
  step_novel(all_nominal_predictors()) %>% #renames "new" hybrids as "new" category so that the model does not return an error
  step_unknown(all_nominal_predictors()) %>% #handling NAs in categorical columns
  step_dummy(all_nominal_predictors()) %>%  #make dummy variables on categorical variables like previous crop
  step_impute_median(all_numeric_predictors())#imputing NA values with median

#DATA PREP
merged_prep <- merged_recipe %>%
  prep()

merged_prep
```

<<<<<<< HEAD
#12. XGBoost
```{r}

```

=======

#12. XGBoost
##a. Setiing Up
```{r, xgb_spec}
###setting up xgboost

xgb_spec <- 
  boost_tree(
    trees = tune(),
    tree_depth = tune(),
    min_n = tune(),
    learn_rate = tune()
    ) %>%
  set_engine ("xgboost") %>%
  set_mode("regression")
  

```

##b. Cross Validation
```{r}
### cross-validation
set.seed(235)
resampling_foldcv <- vfold_cv(merged_train, v = 10)

###Sampling using Latin-Hypercube
xgb_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  learn_rate(),
  trees(),
  size = 50
)
### Adaptive Grid Search to Identify best model later
set.seed(76544)
registerDoParallel(cores = parallel::detectCores() - 1) 
xgb_res <- tune_race_anova(object = xgb_spec,
                      preprocessor = merged_recipe,
                      resamples = resampling_foldcv,
                      grid = xgb_grid,
                      control = control_race(save_pred = TRUE))

stopImplicitCluster()

plot_race(xgb_res)

### Select the best model based on Race Results (highest r2)
best_r2 <- xgb_res %>%
  select_best(metric = "rsq") %>%
  mutate(source = "best_r2")

###Final Specification
final_spec <- boost_tree(
  trees = best_r2$trees,
  tree_depth = best_r2$tree_depth,
  min_n = best_r2$min_n,
  learn_rate = best_r2$learn_rate
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

##c. Validation
```{r}
###Validation
set.seed(10)
final_fit_xgb <- last_fit(final_spec,
                merged_recipe,
                split = merged_split)
final_fit_xgb %>%
  collect_predictions() #provides predictions on the test set
final_fit_xgb %>%
  collect_metrics() #gives the rmse and r2 on the test set
```




>>>>>>> f090c2ee1b60f325a366bfff656a30684f17ef52
#13. LightGBM
```{r, lgbm_spec}
lgbm_spec <- boost_tree(
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  loss_reduction = tune(),
  min_n = tune(),
  sample_size = tune()
) %>%
  set_engine("lightgbm") %>%
  set_mode("regression")
lgbm_spec


#creating workflow
wf <- workflow() %>%
  add_recipe(rec) %>%
  add_model(lgbm_model)
```

```{r}
set.seed(235) 
resampling_foldcv <- vfold_cv(merged_train, # Create 10-fold cross-validation resampling object from training data
                              v = 10)

resampling_foldcv
resampling_foldcv$splits[[1]]


#Hyperparameter grid with Latin Hypercube Sampling
lgbm_grid <- grid_latin_hypercube(
  tree_depth(), 
  min_n(),
  learn_rate(),
  loss_reduction(),
  trees(),
  sample_prop(),
  size = 100
)

lgbm_grid

#plot
ggplot(data = lgbm_grid,
       aes(x = tree_depth, 
           y = min_n)) +
  geom_point(aes(color = factor(learn_rate),
                 size = trees),
             alpha = .5,
             show.legend = FALSE)
```

#Model Tuning
```{r xgb_grid_result}
set.seed(76544)
registerDoParallel(cores = parallel::detectCores() -1)
lgbm_res <- tune_race_anova(object = lgbm_spec,
                      preprocessor = merged_recipe,
                      resamples = resampling_foldcv,
                      grid = lgbm_grid,
                      control = control_race(save_pred = TRUE))
stopImplicitCluster()


beepr::beep()
lgbm_res$.metrics[[2]]
```

```{r}
#Select Best Models based on lowest RMSE
best_rmse <- lgbm_res %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse")

best_rmse

#Select Best Models based on lowers RMSE within 1% loss
best_rmse_pct_loss <- lgbm_res %>% 
  select_by_pct_loss("min_n",
                     metric = "rmse",
                     limit = 1
                     )%>% 
  mutate(source = "best_rmse_pct_loss")

best_rmse_pct_loss

# Based on lowest RMSE within 1 se
best_rmse_one_std_err <- lgbm_res %>% 
  select_by_one_std_err(metric = "rmse",
                        eval_time = 100,
                        trees
                        )%>% 
  mutate(source = "best_rmse_one_std_err")

best_rmse_one_std_err

# Based on greatest R2
best_r2 <- lgbm_res %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2")

best_r2

# Based on lowers R2 within 1% loss
best_r2_pct_loss <- lgbm_res %>% 
  select_by_pct_loss("min_n",
                     metric = "rsq",
                     limit = 1
                     ) %>% 
  mutate(source = "best_r2_pct_loss")

best_r2_pct_loss

# Based on lowest R2 within 1 se
best_r2_one_std_error <- lgbm_res %>% 
  select_by_one_std_err(metric = "rsq",
                        eval_time = 100,
                        trees
                        ) %>%
  mutate(source = "best_r2_one_std_error")

best_r2_one_std_error
```

#Compare and Finalize Model
```{r comparing values}
best_rmse %>% 
  bind_rows(best_rmse_pct_loss, 
            best_rmse_one_std_err, 
            best_r2, 
            best_r2_pct_loss, 
            best_r2_one_std_error)
```

#Final Specification

```{r final_spec_fit}
final_spec <- boost_tree(
  trees = best_r2$trees,           # Number of boosting rounds (trees)
  tree_depth = best_r2$tree_depth, # Maximum depth of each tree
  min_n = best_r2$min_n,           # Minimum number of samples to split a node
  learn_rate = best_r2$learn_rate  # Learning rate (step size shrinkage)
) %>%
  set_engine("lightgbm") %>%
  set_mode("regression")

final_spec
```

#Final Fit and Predictions

#Validation

```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                merged_recipe,
                split = merged_split)

final_fit %>%
  collect_predictions()
```

#Evaluate on Test Set

```{r final_fit_metrics}
final_fit %>%
  collect_metrics()
```


#Evaluate on Training Set

```{r}
final_spec %>%
  fit(yield_mg_ha ~ .,
      data = bake(merged_prep, 
                  merged_train)) %>%
  augment(new_data = bake(merged_prep, 
                          merged_train)) %>% 
  rmse(yield_mg_ha, .pred) %>%
  bind_rows(
    
    
# R2
final_spec %>%
  fit(yield_mg_ha ~ .,
      data = bake(merged_prep, 
                  merged_train)) %>%
  augment(new_data = bake(merged_prep, 
                          merged_train)) %>% 
  rsq(yield_mg_ha, .pred))
```

#14: Comparing Predicted vs Observed Plots

```{r}
final_fit_xgb %>%
  collect_predictions() %>%
  ggplot(aes(x = yield_mg_ha,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

```{r}
final_fit_lgbm %>%
  collect_predictions() %>%
  ggplot(aes(x = yield_mg_ha,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40))
```


#15: Comparing Variable Importances
```{r}
#XGBoost
final_spec_xgb %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, merged_train)) %>% #There little change in variable improtance if you use full dataset
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)


#LightGBM
final_spec_lgbm %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, merged_train)) %>% 
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
```


#16. Comparison of RMSE and R2 Values
```{r}
#Make a table of the RMSE and R2 Values to compare directly
```

#17. Predicting Testing_Submission
```{r}
#Which model we chose
```
