---
title: "Data Science Applied to Ag - Final Project - ML"
format:
  html:
    embed-resources: true
    toc: true
    theme: cerulean
author: Group 1 (Mari, Oluwatomide)
---

```{r}
# shared_lib_path <- "/work/crss8030/instructor_data/shared_R_libs"

#class_packages <- c("tidymodels", "tidyverse", "vip", "ranger", "finetune", "parsnip", "reticulate", "xgboost", "doParallel", "lme4", "here")

#install.packages(class_packages, lib = shared_lib_path)
```

# Locate the installed libraries (important step while running the script in cluster - sapelo2)

```{r}
# .libPaths(c("/work/crss8030/instructor_data/shared_R_libs", .libPaths()))
```

#1. Loading Library
```{r, warning = False, message = False}
#install.packages("lightgbm")
library(tidyverse)
library(lightgbm)
library(xgboost)
library(caret)
library(lubridate)
library(readr)
library(dplyr)
library(tidymodels)
library(vip)
library(ranger)
library(finetune)
library(ggridges)
library(doParallel)
library(glmnet)
library(finetune)
library(janitor)
library(xgboost)
library(lme4)
library(quarto)
library(here)
here::here()
```



#2. Viewing Files
```{r, warning = False, message = False}
training_metadata <- read_csv(here("data", "training", "training_meta.csv"))

training_soil <- read_csv(here("data", "training", "training_soil.csv"))

training_trait <- read_csv(here("data", "training", "training_trait.csv"))

testing_submission <- read_csv(here("data", "testing", "testing_submission.csv"))

testing_meta <- read_csv(here("data", "testing", "testing_meta.csv"))

testing_soil <- read_csv(here("data", "testing", "testing_soil.csv"))
```


#3. Removing Duplicates and determining day of the year
```{r, warning = False, message = False}

training_soil1 <- training_soil %>%
  select(!(year)) %>%
  separate(site,
                   into = c("site", "year")) %>%
  mutate(year = as.numeric(year)) %>%
  filter(nchar(as.character(year)) == 4) 

training_metadata1<- training_metadata %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>% #remove x column
  unique() %>%
  filter(
    latitude >= 24.5 & latitude <= 49.5,
    longitude >= -125 & longitude <= -66.5
  ) ##Removing sites outside US boundaries

write_csv(training_metadata1,
          here::here("data", "training", "training_metadata1.csv"))

training_trait1 <- training_trait %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>% #remove x column with the -Dry, -Early, -Late
  mutate(
    across(c(date_planted, date_harvested),
           ~as.Date(.x, format = "%m/%d/%y")),
    
    day_of_year_planted = yday(date_planted),
    day_of_year_harvested = yday(date_harvested), .before = date_harvested
  )

view(training_soil1)
view(training_metadata1)
view(training_trait1)
```

#4. Creating Merged Dataset
```{r, warning = False, message = False}
training_merged <- training_trait1 %>%
  left_join(training_soil1, by = c("site", "year")) %>%
  left_join(training_metadata1, by = c("site", "year"))

view(training_merged)
```

#5. Mapping Sites
```{r, warning = False, message = False}
library(USAboundaries)
library(ggplot2)
library(dplyr)
library(sf)
site_locations <- us_states() %>%
filter(!(state_abbr %in% c("PR", "AK", "HI")))
ggplot() +
  geom_sf(data = site_locations) +
  geom_point(data = training_merged, 
             aes(x = longitude,
                 y = latitude,
                color = yield_mg_ha)
  )

#Making a map to identify the site distributions. PR, AK, HI were removed since they did not have sites. Points exist in the Mediterranean between Corsica and Italy, i.e. Site NEH2 (which exists in the final testing_submission file).
```


#6. Matching Sites Across Files
```{r, warning = False, message = False}

#Making Sites Only Identical Across Training and Testing

setdiff(testing_submission$site, training_merged$site)  # sites in df1 but not df2
setdiff(training_merged$site, testing_submission$site)  # sites in df2 but not df1

#Drop sites from training_merged that to not need to be predicted to complete testing_submission. ONH1 and 2 are in proximity to one another, however to not appear in the testing_submission dataframe. ONH3 (a new site) appears in testing_submission. Potential to predict ONH3 based on ONH1 and ONH2, so will not remove them.

training_merged1 <- training_merged %>%
  filter(!(site %in% c("IAH1a" ,"IAH1b" ,"IAH1c", "IAH3" , "MOH2" , "NYH1" ,"TXH2" , "IAH1" , "KSH1" , "NEH4" , "SDH1" , "ARH1" , "ARH2",  "COH1" ,"GEH1",  "TXH3" , "TXH4" , "NYS1" ))) %>%
  mutate(id = row_number(), .before = "year") #adding an ID column at the front of the table to identify rows

write_csv(training_merged1,
          here::here("data", "training", "training_merged1.csv"))

#confirm: 
setdiff(testing_submission$site, training_merged1$site)  # sites in df1 but not df2
setdiff(training_merged1$site, testing_submission$site)  # sites in df2 but not df1

```

```{r, warning = False, message = False}
fieldweather <- read_csv(here("data", "fieldweatherdata.csv"))

fieldweatherdata <- fieldweather %>%
  select(-gdd)

summary(fieldweatherdata)

```


#7. Density plots  
```{r}
#To explore the weather data distributions

fieldweatherdata %>% 
  pivot_longer(cols = dayl_s:gdd_rounded) %>%
  ggplot(aes(x = value)) +
  geom_density() + 
  facet_wrap(~name, scales = "free")

```

```{r}
fieldweather1 <- fieldweatherdata %>%
  # Selecting needed variables
  dplyr::select(-tile, -altitude) %>%
# Creating a date class variable  
  mutate(date_chr = paste0(year, "/", yday)) %>% #paste0 helps to merge columns together
  mutate(date = as.Date(date_chr, "%Y/%j"))

```


```{r, warning=FALSE}
weather_features <- training_merged1 %>%
  select(site, year, date_planted, date_harvested) %>%
  distinct() %>%
  left_join(fieldweather1, by = c("site", "year")) %>%
  filter(date >= date_planted,
         date <= date_harvested) %>%
  group_by(site, year, date_planted, date_harvested) %>%
  summarise(
    n_days = n(),
    sum_gdd = sum(gdd_rounded, na.rm = TRUE),
    across(
      c(dayl_s, srad_w_m_2, tmax_deg_c, tmin_deg_c, vp_pa),
      mean, na.rm = TRUE,
      .names = "mean_{.col}"
    ),
    sum_precip = sum(prcp_mm_day, na.rm = TRUE),
    .groups = "drop"
  )

weather_merged <- training_merged1 %>%
  filter(year != 2014) %>%
  left_join(weather_features,
            by = c("site", "year", "date_planted", "date_harvested"))


weather_merged
write_csv(weather_merged,
          here::here("output", "weather_merged.csv"))
```

# 8. Data Visualization 
```{r}
#Plotting all variables by site

allplots <- weather_merged %>%
  pivot_longer(sum_gdd:sum_precip) %>%
  group_by(name) %>%
  nest() %>%
  mutate(plot = map2(data, name,
                     ~ggplot(data = .x,
       aes(x = value,
           y = site,
           fill = stat(x))
       ) +
  geom_density_ridges_gradient(scale = 3,
                               rel_min_height = 0.01) +
  scale_fill_viridis_c(option = "C", alpha = .8) +
  theme(legend.position = "none") +
  labs(x= .y,
       y = "site")))
  
allplots  
```

```{r}
allplots$plot
```

#9. Pre-Processing Before MLA
```{r}
set.seed(76332)

# Splitting the Data for Training
merged_split <- initial_split(weather_merged, prop = 0.7, strata = yield_mg_ha)
merged_split

merged_train <- training(merged_split) %>%
  select(-c(date_planted, date_harvested))
merged_train

merged_test <- testing(merged_split)
merged_test
```

#10. Distribution of Yield
```{r, density plot}
density_plot <- ggplot() +
  geom_density(data = merged_train, 
               aes(x = yield_mg_ha),
               color = "red") +
  geom_density(data = merged_test, 
               aes(x = yield_mg_ha),
               color = "blue") 

cat(paste0("Saving...... density plot of train and test data"))

ggsave(plot = density_plot, 
       path = here("output", "png"),
       filename = "density_plot_test_train.png",
       height = 6,
       width = 9,
       dpi = 600)
```

#11. Recipe and Prep for Training Model
```{r}
#RECIPE

cat("\nCreate recipe for data preprocessing\n")

merged_recipe <- recipe(yield_mg_ha ~ ., data = merged_train) %>%
  step_novel(all_nominal_predictors()) %>% #renames "new" hybrids as "new" category so that the model does not return an error
  step_unknown(all_nominal_predictors()) %>% #handling NAs in categorical columns
  step_dummy(all_nominal_predictors()) %>%  #make dummy variables on categorical variables like previous crop
  step_impute_median(all_numeric_predictors())#imputing NA values with median

merged_recipe

#DATA PREP
merged_prep <- merged_recipe %>%
  prep()

merged_prep
```

```{r}
# helps to see all the available hyperparameter we can tune for a model
show_model_info("boost_tree")
# helps to see all the models that we can run with the package parsnip
get_from_env("models")
```

#12. XGBoost
##a. Setting Up
```{r, xgb_spec}
###setting up xgboost

xgb_spec <- 
  boost_tree(
    trees = tune(),
    tree_depth = tune(),
    min_n = tune(),
    learn_rate = tune()
    ) %>%
  set_engine ("xgboost") %>%
  set_mode("regression")

xgb_spec
```

##b. Cross Validation
```{r}
### cross-validation
set.seed(235)
resampling_foldcv <- vfold_cv(merged_train, v = 10)

# Create leave one year out cv object from the sampling data
resampling_fold_loyo <- group_vfold_cv(merged_train,
                                       group = year)

# Create leave one location out cv object from the sampling data
resampling_fold_loso <- group_vfold_cv(merged_train,
                                       group = site)
###Sampling using Latin-Hypercube
set.seed(12345)
xgb_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  learn_rate(),
  trees(),
  size = 50
)
xgb_grid

ggplot(data = xgb_grid,
       aes(x = tree_depth, 
           y = min_n)) +
  geom_point(aes(color = factor(learn_rate),
                 size = trees),
             alpha = .5,
             show.legend = FALSE)

### Adaptive Grid Search to Identify best model later
n_cores <- as.numeric(Sys.getenv("SLURM_CPUS_ON_NODE"))

if (is.na(n_cores)) n_cores <- parallel::detectCores() - 1

# Start the Cluster
cl <- makePSOCKcluster(n_cores)

registerDoParallel(cl)

cat(paste0("\nFound and registered ", n_cores, " cores to work with\n"))
```

```{r}
set.seed(76544)

# Create the list of CV techniques so we can loop through them
cv_list <- list(
  vfold = resampling_foldcv,
  year = resampling_fold_loyo,
  site = resampling_fold_loso
)

# Create a empty list to store the results from the loop
results <- list()

# Create our loop
for (i in seq_along(cv_list)) {
  name <- names(cv_list)[i]
  results[[name]] <- tune_race_anova(object = xgb_spec,
                                     preprocessor = merged_recipe,
                                     resamples = cv_list[[i]],
                                     grid = xgb_grid,
                                     control = control_race(save_pred = TRUE,
                                                            parallel_over = "everything"))
}

stopCluster(cl)


# Create a dataframe structure from the list obtained after running the loop
results_df <- tibble(method = names(results),
                     diff_cv = results) %>%  
  # Collect the metrices for each CV techniques using map function
  mutate (metrices = map2(diff_cv, method,
                          ~.x %>% 
                            collect_metrics() %>%
                            mutate(method = .y, .before = "trees")))
results_df$metrices[[1]]

# bind all the metrices together so to select the best performing one
all_metrices <- do.call(bind_rows, results_df$metrices)


# Automating to pull the best method out of 3 we ran
best_method <- all_metrices %>%
  filter(.metric == "rmse") %>%
  slice_min(mean, n = 1) %>%
  pull(method)

# Getting the metrice (hyperparameter values of the best performing CV)
best_cv_object <- results_df %>%
  filter(method == best_method) %>%
  pull(diff_cv) %>% 
  first()
  
```

```{r}
### Select the best model based on Race Results (highest rmse)
best_rmse_xgb <- best_cv_object %>%
  select_best(metric = "rmse") %>%
  mutate(source = "best_rmse")
best_rmse_xgb


### Select the best model based on Race Results (highest r2)
best_r2_xgb <- best_cv_object %>%
  select_best(metric = "rmse") %>%
  mutate(source = "best_r2")
best_r2_xgb


## Compare and Finalize Model
best_rmse_xgb %>% 
  bind_rows(best_rmse_xgb, 
            best_r2_xgb) %>%
  dplyr::select(source, everything())

###Final Specification
final_spec_xgb <- boost_tree(
  trees = best_r2_xgb$trees,
  tree_depth = best_r2_xgb$tree_depth,
  min_n = best_r2_xgb$min_n,
  learn_rate = best_r2_xgb$learn_rate
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")

final_spec_xgb
```

##. Final Fit and Predictions
```{r, warning=FALSE}
###Validation
set.seed(10)
final_fit_xgb <- last_fit(final_spec_xgb,
                merged_recipe,
                split = merged_split)
final_fit_xgb %>%
  collect_predictions() #provides predictions on the test set
final_fit_xgb %>%
  collect_metrics() %>% #gives the rmse and r2 on the test set
  mutate(estimate = round(.estimate, 3))
```

#13. LightGBM
```{r, lgbm_spec}
###setting up xgboost

lgbm_spec <- 
  boost_tree(
    trees = tune(),
    tree_depth = tune(),
    min_n = tune(),
    learn_rate = tune()
    ) %>%
  set_engine ("lightgbm") %>%
  set_mode("regression")

lgbm_spec
```

##b. Cross Validation
```{r}
### cross-validation
set.seed(1235)
resampling_foldcv <- vfold_cv(merged_train, v = 10)

# Create leave one year out cv object from the sampling data
resampling_fold_loyo <- group_vfold_cv(merged_train,
                                       group = year)

# Create leave one location out cv object from the sampling data
resampling_fold_loso <- group_vfold_cv(weather_train,
                                       group = site)
###Sampling using Latin-Hypercube
set.seed(123456)
lgbm_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  learn_rate(),
  trees(),
  size = 50
)
lgbm_grid

ggplot(data = lgbm_grid,
       aes(x = tree_depth, 
           y = min_n)) +
  geom_point(aes(color = factor(learn_rate),
                 size = trees),
             alpha = .5,
             show.legend = FALSE)

### Adaptive Grid Search to Identify best model later
n_cores <- as.numeric(Sys.getenv("SLURM_CPUS_ON_NODE"))

if (is.na(n_cores)) n_cores <- parallel::detectCores() - 1

# Start the Cluster
cl <- makePSOCKcluster(n_cores)

registerDoParallel(cl)

cat(paste0("\nFound and registered ", n_cores, " cores to work with\n"))
```

```{r}
set.seed(176544)

# Create the list of CV techniques so we can loop through them
cv_list1 <- list(
  vfold = resampling_foldcv,
  year = resampling_fold_loyo,
  site = resampling_fold_loso
)

# Create a empty list to store the results from the loop
results1 <- list()

# Create our loop
for (i in seq_along(cv_list)) {
  name <- names(cv_list)[i]
  results1[[name]] <- tune_race_anova(object = lgbm_spec,
                                     preprocessor = merged_recipe,
                                     resamples = cv_list1[[i]],
                                     grid = lgbm_grid,
                                     control = control_race(save_pred = TRUE,
                                                            parallel_over = "everything"))
}

stopCluster(cl)


# Create a dataframe structure from the list obtained after running the loop
results_df1 <- tibble(method = names(results1),
                     diff_cv = results1) %>%  
  # Collect the metrices for each CV techniques using map function
  mutate (metrices = map2(diff_cv, method,
                          ~.x %>% 
                            collect_metrics() %>%
                            mutate(method = .y, .before = "trees")))
results_df1$metrices[[1]]

# bind all the metrices together so to select the best performing one
all_metrices1 <- do.call(bind_rows, results_df1$metrices)


# Automating to pull the best method out of 3 we ran
best_method1 <- all_metrices1 %>%
  filter(.metric == "rmse") %>%
  slice_min(mean, n = 1) %>%
  pull(method)

# Getting the metrice (hyperparameter values of the best performing CV)
best_cv_object1 <- results_df1 %>%
  filter(method == best_method1) %>%
  pull(diff_cv) %>% 
  first()
  
```

```{r}
### Select the best model based on Race Results (highest rmse)
best_rmse_lgbm <- best_cv_object1 %>%
  select_best(metric = "rmse") %>%
  mutate(source = "best_rmse")
best_rmse_lgbm


### Select the best model based on Race Results (highest r2)
best_r2_lgbm <- best_cv_object1 %>%
  select_best(metric = "rmse") %>%
  mutate(source = "best_r2")
best_r2_lgbm


## Compare and Finalize Model
best_rmse_lgbm %>% 
  bind_rows(best_rmse_lgbm, 
            best_r2_lgbm) %>%
  dplyr::select(source, everything())

###Final Specification
final_spec_lgbm <- boost_tree(
  trees = best_r2_lgbm$trees,
  tree_depth = best_r2_lgbm$tree_depth,
  min_n = best_r2_lgbm$min_n,
  learn_rate = best_r2_lgbm$learn_rate
) %>%
  set_engine("lightgbm") %>%
  set_mode("regression")

final_spec_lgbm
```

##. Final Fit and Predictions
```{r, warning=FALSE}
###Validation
set.seed(1000)
final_fit_lgbm <- last_fit(final_spec_lgbm,
                merged_recipe,
                split = merged_split)
final_fit_lgbm %>%
  collect_predictions() #provides predictions on the test set
final_fit_lgbm %>%
  collect_metrics() %>% #gives the rmse and r2 on the test set
  mutate(estimate = round(.estimate, 3))
```

#Final Specification

```{r final_spec_fit}
final_spec_lgbm <- boost_tree(
  trees = best_r2_lgbm$trees,           # Number of boosting rounds (trees)
  tree_depth = best_r2_lgbm$tree_depth, # Maximum depth of each tree
  min_n = best_r2_lgbm$min_n,           # Minimum number of samples to split a node
  learn_rate = best_r2_lgbm$learn_rate  # Learning rate (step size shrinkage)
) %>%
  set_engine("lightgbm") %>%
  set_mode("regression")

final_spec_lgbm
```

#Final Fit and Predictions

#Validation

```{r final_fit}
set.seed(10)
final_fit_lgbm <- last_fit(final_spec_lgbm,
                merged_recipe,
                split = merged_split)

final_fit_lgbm %>%
  collect_predictions()
```

#Evaluate on Test Set

```{r final_fit_metrics}
final_fit_lgbm %>%
  collect_metrics()
```


#Evaluate on Training Set

```{r}
final_spec_lgbm %>%
  fit(yield_mg_ha ~ .,
      data = bake(merged_prep, 
                  merged_train)) %>%
  augment(new_data = bake(merged_prep, 
                          merged_train)) %>% 
  rmse(yield_mg_ha, .pred) %>%
  bind_rows(
    
    
# R2
final_spec_lgbm %>%
  fit(yield_mg_ha ~ .,
      data = bake(merged_prep, 
                  merged_train)) %>%
  augment(new_data = bake(merged_prep, 
                          merged_train)) %>% 
  rsq(yield_mg_ha, .pred))
```

#14: Comparing Predicted vs Observed Plots

```{r}
Publication_ready_xgb <- final_fit_xgb %>%
  collect_predictions() %>%
  ggplot(aes(x = yield_mg_ha,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 

ggsave(plot = publication_ready_xgb, 
       path = here("output", "png"),
       filename = "model_perf_test_data.png",
       height = 6,
       width = 9,
       dpi = 600)
```

```{r}
Publication_ready_lgbm <- final_fit_lgbm %>%
  collect_predictions() %>%
  ggplot(aes(x = yield_mg_ha,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40))

ggsave(plot = publication_ready_lgbm, 
       path = here("output", "png"),
       filename = "model_perf_test_data.png",
       height = 6,
       width = 9,
       dpi = 600)
```


#15: Comparing Variable Importances
```{r}
#XGBoost
vip_xgb <- final_spec_xgb %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, merged_train)) %>% 
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL) +
  theme(panel.background  = element_rect(fill = "gray82"),
        panel.grid = element_blank())

cat(paste0("\nSaving...... publication ready plot for variable importance\n"))

ggsave(plot = vip_xgb, 
       path = here("output/png"),
       filename = "vip_xgb_test_data.png",
       height = 6,
       width = 9,
       dpi = 600)


#LightGBM
vip_lgbm <- final_spec_lgbm %>%
  fit(yield_mg_ha ~ .,
         data = bake(merged_prep, merged_train)) %>% 
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL) +
  theme(panel.background  = element_rect(fill = "gray82"),
        panel.grid = element_blank())

cat(paste0("\nSaving...... publication ready plot for variable importance\n"))

ggsave(plot = vip_lgbm, 
       path = here("output/png"),
       filename = "vip_lgbm_test_data.png",
       height = 6,
       width = 9,
       dpi = 600)
```


#16. Comparison of RMSE and R2 Values
```{r}
#Make a table of the RMSE and R2 Values to compare directly

metrics_xgb <- final_fit_xgb %>%
  collect_metrics()

metrics_lgbm <- final_fit_lgbm %>%
  collect_metrics()


write_csv(metrics_xgb,
          here::here("output", "xgb_metrics.csv"))

write_csv(metrics_xgb,
          here::here("output", "lgbm_metrics.csv"))

```

```{r}
metrics_xgb <- final_fit_xgb %>%
  collect_metrics() %>%
  dplyr::mutate(model = "xgboost")

metrics_lgbm <- final_fit_lgbm %>%
  collect_metrics() %>%
  dplyr::mutate(model = "lightgbm")

all_metrics <- dplyr::bind_rows(metrics_xgb, metrics_lgbm)

comparison_table <- all_metrics %>%
  dplyr::filter(.metric %in% c("rmse", "rsq")) %>%
  tidyr::pivot_wider(names_from = .metric, values_from = .estimate)

comparison_table

write_csv(comparison_table,
          here::here("output", "model_comparison_metrics.csv"))

```



#17. Predicting Testing_Submission
```{r}
#Wrangling testing soil to separate year from site in the site column before merging for prediction
testing_soil1 <- testing_soil %>%
  separate(site, 
                  into = c("site", "x")) %>%
  select(!(x)) %>%
  unique()

merged_submission <- testing_submission %>%
  left_join(testing_soil1, by = c("site", "year")) %>%
  left_join(testing_meta, by = c("site", "year")) 
view(merged_submission)

#selecting model with best r2
comparison <- all_metrics %>%
  dplyr::filter(.metric == "rsq")

best_model <- comparison %>%
  dplyr::slice_max(.estimate, n = 1) %>%
  dplyr::pull(model)

#selecting the best model to predict based on r2
best_fit <- if (best_model == "xgboost") {
  final_fit_xgb
} else {
  final_fit_lgbm
}

best_workflow <- best_fit %>%
  extract_workflow()

# predicting the yield for all rows
pred_yield <- predict(best_workflow, merged_submission) %>%
  dplyr::pull(.pred)

# replacing all NA values with predicted value
merged_submission_final <- merged_submission %>%
  dplyr::mutate(
    yield_mg_ha = dplyr::if_else(
      is.na(yield_mg_ha),
      pred_yield,
      yield_mg_ha
    )
  ) %>% 
  dplyr::select(year, site, hybrid, yield_mg_ha)

write_csv(merged_submission_final,
          here::here("output", "testing_submission.csv"))
```


```{r}
#knitr::purl("project_code_teach.qmd", output = "project_code_teach1.R", documentation = 0)
```
